---
title: "EDA Using R/RStudio"
author: "Stat 230"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Instructions

1. Work through the [EDA examples in the R manual](https://stat230.github.io/r-manual-kuiper/01-eda.html)

2. Explore the `HELPrct` data set to try to find some interesting insights. 
You can find details (variable names, etc.) about the data set in the [online codebook](https://www.rdocumentation.org/packages/mosaicData/versions/0.20.2/topics/HELPrct). Explore the data set using plots, tables, and other numeric summaries. 
Select three of your favorite plots and tell a story (in writing) about each of them. 

Be sure to include this code chunk so that the data and necessary R packages
are available in this document.

```{r message=FALSE}
# load the data set from the url
HELPrct <- read.csv("https://aloy.rbind.io/data/HELPrct.csv")

# loading the necessary packages
library(ggformula)
library(mosaic)
```

Feel free to add code chunks below as you work through the R Manual, but remember that
your goal is to create three plots that tell a story.


## PLOT 1

```{r}
# put the code for your plot here
```

(Include the description for your plot here.)


## PLOT 2

```{r}
# put the code for your plot here
```
(Include the description for your plot here.)


## PLOT 3

```{r}
# put the code for your plot here
```
(Include the description for your plot here.)