---
title: "Models to Compare Two Population Means"
date: "2022-09-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## Part 1: Understanding the study design


### 1.

Write your answer here...

### 2.

Write your answer here...

### 3.

Write your answer here...

### 4.

```{r plot}
# Delete this line and put your code here
```

Write your interpretation here...



## Part 2: Statistical models for the two-sample t-test

### Task 2. 

```{r sample-means}
# Delete this line and put your code here
```



## Part 3: Checking assumptions


### 7.

```{r reisdual-histogram}
# Delete this line and put your code here
```

Write your interpretation here...

### 8.

```{r equal-variances}
# Delete this line and put your code here
```

Write your interpretation here...

### 9.

```{r resid-StudentID}
# Delete this line and put your code here
```

Write your interpretation here...

### 10.

```{r t-test}
# Delete this line and put your code here
```

Write your interpretation here...



## Part 4: Fitting a regression model to compare means



### 11.

```{r slr-model}
# Delete this line and put your code here
```


### 12.

```{r regression-inference}
# Delete this line and put your code here
```

Write your interpretation here...

### 13.

```{r relevel-indicator}
# Delete this line and put your code here
```

Write your interpretation here...



## Part 5: Check the model 


### 14.

```{r resid-plots}
# Delete this line and put your code here
```

Write your interpretation here...




## Part 6: Communicate the results

### 15.

```{r plot-model}
# Delete this line and put your code here
```

Write your interpretation here...

