---
title: "Models to Compare Two Population Means"
date: "Winter 2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


## Part 1: Understanding the study design


### 1.

Write your answer here...

### 2.

Write your answer here...

### 3.

Write your answer here...

### 4.

```{r plot}
# Delete this line and put your code here
```

Write your interpretation here...



## Part 2: Fitting a regression model to compare means

### 11.

```{r slr-model}
# Delete this line and put your code here
```


### 12.

```{r regression-inference}
# Delete this line and put your code here
```

Write your interpretation here...

### 13.

```{r relevel-indicator}
# Delete this line and put your code here
```

Write your interpretation here...



## Part 3: Check the model


### 14.

```{r resid-plots}
# Delete this line and put your code here
```

Write your interpretation here...




## Part 4: Communicate the results

### 15.

```{r plot-model}
# Delete this line and put your code here
```

Write your interpretation here...


## Part 5: Comparing the results to the two-sample t-test

### Task 5.

Write your answer here...