---
output: pdf_document
---

# Post-Hoc Analysis for One-Way ANOVA

## Names:

### Stat 230, Winter 2024

```{r setup, include=FALSE}
library(mosaic)
library(ggformula)
```

The data set can be loaded using the code

```{r}
corn <- read.csv("https://aloy.rbind.io/data/cornplot.csv")
```

## EDA

To begin, explore the data set by completing the following tasks:

### Task 1.

Create an individual values plot of the data. Comment on the distribution within each treatment group. Are there outliers? Are the distributions skewed? Are there unequal variances? 


### Task 2.

Reflecting on your answers to task 1, are you worried about any of the conditions necessary for one-way ANOVA?

## ANOVA

Now, let's use the one-way ANOVA model to conduct a first analysis.

### Task 3. 

Use R to fit the one-way ANOVA model for this experiment. Remember that we use the `aov()` function to fit the model and the `summary()` function to print the ANOVA table.

```{r}
# Fill in the blanks to fit the ANOVA model
corn_anova <- aov(___ ~ factor(___), data = ___)
summary(___)
```



### Task 4. 

The ANOVA model predicts that every point in a treatment group will be at the group mean. Use the below code to create a plot of the fitted ANOVA model. Comment on the fit of this model to your data.

```{r}
group_means <- corn |> 
  group_by(Treatment) |> 
  summarize(avg = mean(Weight))

gf_jitter(Weight ~ Treatment, data = corn, width= 0.1) |>
  gf_summary(fun.y = "mean", color = "darkorange", size = 3, geom = "point")
```


### Task 5. 

Construct a plot of the residuals vs. the explanatory variable (Treatment) and a histogram of the residuals for the ANOVA model. Comment on the fit of the ANOVA model.


## Post-hoc analysis



### Task 6.  

Calculate 95% confidence intervals for the 10 pairwise comparisons of each of the treatments using the `TukeyHSD()` command.

```{r}
TukeyHSD(___, conf.level = ___)
```



### Task 7.

Summarize, broadly, what these intervals tell you about how the treatments compare to each other.


