---
title: "Model Diagnostics and Remedial Measures"
date: "Winter 2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the require R packages in this setup chunk
library(ggformula)
library(mosaic)

# Load the data set in the set up chunk
games1 <- read.csv("https://aloy.rbind.io/kuiper_data/Games1.csv")
```

### Activity A

#### a.

Write your answer here...

#### b.

Write your answer here...

#### c.

Write your answer here...

#### d.

Write your answer here...

#### e.

Write your answer here...




### Exercise E.10

Here is code to calculate the residuals from the one-way ANOVA model.

```{r}
games_anova <- aov(Time ~ Type, data = games1)
games_resid <- resid(games_anova)
```


#### a.

```{r e10a}
# Delete this comment and put your code here
```

Write your answer here...


#### b.

```{r e10b}
# Delete this comment and put your code here
```

Write your answer here...


#### c.


Write your answer here...



### Extended Activity 31

```{r}
emissions <- read.csv("https://aloy.rbind.io/kuiper_data/Emissions.csv")
```


#### d.

```{r 31d}
# Delete this comment and put your code here
```

Write your answer here...


#### e.

```{r 31e}
# Delete this comment and put your code here
```

Write your answer here...



### Activity B

This code just creates the `y1` and `y2` samples you'll use in the activity.
```{r}
sub_emissions <- emissions |>
     filter(Year == "1963-1967" | Year == "1970-1971")
y1 <- sub_emissions |> 
  filter(Year == "1963-1967") |> 
  pull(Emissions)
y2 <- sub_emissions |> 
  filter(Year == "1970-1971") |>
  pull(Emissions)
```

#### a.

```{r ActB-a}
# Delete this comment and put your code here
```

#### b.

```{r ActB-b}
# Delete this comment and put your code here
```

#### c.

```{r ActB-c}
# Delete this comment and put your code here
```

Write your answer here...

#### d.

```{r ActB-d}
# Delete this comment and put your code here
```

Write your answer here...

#### e.

```{r ActB-e}
# Delete this comment and put your code here
```

Write your answer here...

#### f.

Write your answer here...

