---
title: "Model Diagnostics and Remedial Measures"
date: "2022-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the require R packages in this setup chunk
library(ggformula)
library(mosaic)

# Load the data set in the set up chunk
games1 <- read.csv("https://aloy.rbind.io/kuiper_data/Games1.csv")
```

### Activity A

#### a.

Write your answer here...

#### b.

Write your answer here...

#### c.

Write your answer here...

#### d.

Write your answer here...

#### e.

Write your answer here...


### Activity 29

```{r}
normal <- read.csv("https://aloy.rbind.io/kuiper_data/Normal.csv")
```

#### a.

```{r 29a}
# Delete this comment and put your code here
```

Write your answer here...


#### b.

```{r 29b}
# Delete this comment and put your code here
```

Write your answer here...


#### c.

```{r 29c}
# Delete this comment and put your code here
```

Write your answer here...


### Exercise E.10

#### a.

```{r e10a}
# Delete this comment and put your code here
```

Write your answer here...


#### b.

```{r e10b}
# Delete this comment and put your code here
```

Write your answer here...


#### c.

```{r e10c}
# Delete this comment and put your code here
```

Write your answer here...



### Extended Activity 31

```{r}
emissions <- read.csv("https://aloy.rbind.io/kuiper_data/Emissions.csv")

sub_emissions <- emissions |> 
  filter(Year == "1963-1967" | Year == "1970-1971")
```


#### c.

```{r 31c}
# Delete this comment and put your code here
```

Write your answer here...



#### d.

```{r 31d}
# Delete this comment and put your code here
```

Write your answer here...



### Activity B

```{r}
y1 <- sub_emissions |> 
  filter(Year == "1963-1967") |> 
  pull(Emissions)
y2 <- sub_emissions |> 
  filter(Year == "1970-1971") |>
  pull(Emissions)
```

#### a.

```{r ActB-a}
# Delete this comment and put your code here
```

#### b.

```{r ActB-b}
# Delete this comment and put your code here
```

#### c.

```{r ActB-c}
# Delete this comment and put your code here
```

Write your answer here...

#### d.

```{r ActB-d}
# Delete this comment and put your code here
```

Write your answer here...

#### e.

```{r ActB-e}
# Delete this comment and put your code here
```

Write your answer here...

#### f.

Write your answer here...




### Activity C

#### Transforming y1

```{r ActC-y1}
# Delete this comment and put your code here
```

#### Transforming y2

```{r ActC-y2}
# Delete this comment and put your code here
```

#### Transforming y3

```{r ActC-y3}
# Delete this comment and put your code here
```
