---
title: "Prediction for Simple Linear Regression"
date: "2022-09-26"
output: html_document
---

```{r setup, include=FALSE}
# Load the required R packages in this setup chunk
library(mosaic)
library(ggformula)

# Load the data set in the set up chunk
cars <- read.csv("https://aloy.rbind.io/kuiper_data/Cars.csv")
```

## Task 1

```{r scatterplot}
# Delete this comment and put your code here
```

Write your answer here...


## Task 2

```{r fit-model}
# Delete this comment and put your code here
```

Write your answer here...


## Task 3

Write your interpretations here...


## Task 4

Write your interpretation here...


## Task 5 

Write your answer here...


## Task 6

```{r post-hoc-intervals, error=FALSE}
# Fill in the blank with your scatterplot code
___ |> gf_lm()
```


## Task 7

```{r fit-model}
# Remember R can be a big calculator
```

Write your answer here...



## Task 8

Write your examples here...


## Task 9


Write your answer here...


## Task 10

Write your answer here...


## Task 11

```{r confidence-interval}
# Delete this comment and put your code here
```

Write your interpretation here...


## Task 12

```{r prediction-interval}
# Delete this comment and put your code here
```

Write your interpretation here...


## Task 13

```{r}
# Fill in the blank with your scatterplot code
___ |> 
  gf_lm(interval = "prediction") |>
  gf_lm(interval = "confidence", alpha = 0.6)
```



## Task 14

```{r diagnostics}
# Delete this comment and put your code here
```

Write your comments here...
