---
title: "Prediction for Simple Linear Regression"
date: "Winter 2024"
output: pdf_document
---

```{r setup, include=FALSE}
# Load the required R packages in this setup chunk
library(mosaic)
library(ggformula)

# Load the data set in the set up chunk
cars <- read.csv("https://aloy.rbind.io/kuiper_data/Cars.csv")
```


# Part I

Here is code to recreate figure 3.1 from the textbook:

```{r warning=FALSE}
gf_point(Price ~ Mileage, data = cars) |> # create the scatterplot, and then
  gf_lm()                                 # add the fitted regression line
```


## Task 1

Fit the model and print the summary, then identify the required parts of the summary (no need to show that you identified them here.)

```{r fit-model}
# Delete this comment and put your code here
```


## Task 2

Write your interpretations here...


## Task 3

Write your interpretation here...


## Task 4

Write your answer here...


## Task 5

Write your answer here...


# Part II

## Task 6

```{r confidence-interval}
# Delete this comment and put your code here
```

Write your interpretation here...


## Task 7

```{r prediction-interval}
# Delete this comment and put your code here
```

Write your interpretation here...


## Task 8

```{r plot-pred-intervals}
# Delete this comment and put your code here
```

## Task 9

Write your answer here...

```{r diagnostics}
# Delete this comment and put your code here
```

Write your comments here...
